Resolvable Ambiguity

### 2019-07-05 paper/paper.tex
Case studies introduction

\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  basewidth={.5em,.5em},
}
\newcommand{\ocaml}{\lstinline[language={[objective]caml}]}

@@

>>>>

\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  basewidth={.5em,.5em},
}
\lstdefinelanguage{syncon}
  {morekeywords={syncon,infix,postfix,prefix,left,right,comment,precedence,forbid,except,type,builtin,rec,grouping},
   sensitive=true,
   morecomment=[l]{//},
   morecomment=[s]{/*}{*/},
   morecomment=[s][<.][.>],
   morestring=[b]",
  }
\newcommand{\ocaml}{\lstinline[language={[objective]caml}]}
\newcommand{\syncon}{\lstinline[language=syncon]}

@@

We have implemented a tool and a small DSL for syntax definition that generates a language definition in the style of Section~\ref{sec:parse-time-disambiguation}, which the tool then uses to construct a parser and do a dynamic resolvability analysis. The tool is implemented in Haskell and uses an off-the-shelf implementation\footnote{\url{http://hackage.haskell.org/package/Earley}} of the Earley parsing algorithm \cite{earleyEfficientContextfreeParsing1970}. The purpose of these case studies is to showcase possible use cases of resolvable ambiguity in the context of a pre-existing general purpose programming language (OCaml, Section~\ref{sec:evaluation-ocaml}), a pre-existing DSL (ORC, Section~\ref{sec:evaluation-orc}), and a new DSL (Cyphym, Section~\ref{sec:evaluation-cyphym}).

The syntax definition DSL is a continuation of the ``syncons'' introduced by \citet{palmkvistCreatingDomainSpecificLanguages2019}, wherein each production is defined separately from each other (one \syncon{syncon} each), and marks are primarily introduced through precedence lists. For example, the running example used in Section~\ref{sec:parse-time-disambiguation} (Fig.~\ref{fig:running-example-definition} on page~\pageref{fig:running-example-definition}) can be written as follows:

\begin{tabular}{ll}
\begin{lstlisting}[language=syncon,boxpos=t]
type Exp
grouping "(" Exp ")"
precedence {
  mul; // higher in precedence list
  add; // means higher precedence
}
\end{lstlisting}
&
\begin{lstlisting}[language=syncon,boxpos=t]
syncon list: Exp =
  "[" (head:Exp (";" tail:Exp))? "]"
infix add: Exp = "+"
infix mul: Exp = "*"
syncon literal: Exp = n:Integer
\end{lstlisting}
\end{tabular}

\noindent The design of the DSL diverges from the formalism in Section~\ref{sec:parse-time-disambiguation} partially for familiarity (precedence instead of marks), but also for composability reasons: the DSL allows marks to be introduced without altering the original definition of a syncon. If marks are needed in non-precedence related situations, e.g., adding sequential composition to the running example, they can be introduced with ``forbids'':

\begin{lstlisting}[language=syncon]
infix seqComp: Exp = ";"
forbid list.head = seqComp
forbid list.tail = seqComp
\end{lstlisting}
